---
title: "6372 Project 2"
author: "Chad Kwong"
date: '2022-07-15'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(glmnet)
library(MASS)
library(tidyverse)
```

# Initial Data loading and approach on Missing values
## Initial Data load in and column manipulation
```{r}
#Combining provided Test and Train sets into one data set for custom training test splits
#salaryDataTrain = read.csv()
#salaryDataTest = read.csv(file.choose(), header = FALSE)
#salaryDataTest = salaryDataTest[-1,]
#colnames(salaryDataTrain) <- c('age', 'workclass', 'fnlwgt', 'education', 'education_num', 'marital_status', 'occupation', 'relationship', 'race', 'sex', 'capital_gain', 'capital_loss', 'hours_per_week', 'native_country', 'money')
#colnames(salaryDataTest)<- colnames(salaryDataTrain)
#salaryData <- rbind(salaryDataTrain, salaryDataTest)
# export dataset to new file
#write.csv(salaryData, file = "/Users/chadkwong/Desktop/CombinedData.csv")

```

```{r}
salaryData = read.csv("data/Dataset/CombinedData.csv")
str(salaryData)
```


```{r}
#converting variables to factors, making sure that levels are equal, and changing ? to NA
salaryData[salaryData==" ?"] = NA
salaryData$money[salaryData$money==' <=50K.'] = ' <=50K'
salaryData$money[salaryData$money==' >50K.'] = ' >50K'
salaryData$age = as.numeric(salaryData$age)
salaryData$workclass=as.factor(salaryData$workclass)
salaryData$education=factor(salaryData$education)
salaryData$marital_status=factor(salaryData$marital_status)
salaryData$occupation=factor(salaryData$occupation)
salaryData$relationship=factor(salaryData$relationship)
salaryData$race=factor(salaryData$race)
salaryData$sex=factor(salaryData$sex)
salaryData$native_country=factor(salaryData$native_country)
salaryData$money=factor(salaryData$money)
salaryData = salaryData[,-1]
```

## Missing Value Investigation
```{r}
#Initial table of missing values
table(is.na(salaryData))
#Gathering all na values and creating an organizational dataframe to display where each missing value is
#create data frame with 0 rows and 4 columns
naCol <- data.frame(matrix(ncol = 5, nrow = 0))
#provide column names
colnames(naCol) <- c('varName', 'varType', 'varID#', 'naCount', 'PercentageNA')
for(i in 1:ncol(salaryData)){
  if(length(table(is.na(salaryData)))==2){
    naCol[i,] = c("dummy","dummy",i,table(is.na(salaryData[,i]))[2],"dummy")
  }
}
naCol = na.omit(naCol)
naCol$`varID#` = as.integer(naCol$`varID#`)
naCol$naCount = as.integer(naCol$naCount)
naCol$PercentageNA = round(naCol$naCount/length(salaryData$age),4)*100
for(i in 1:length(naCol$varName)){
  naCol$varName[i] = colnames(salaryData)[naCol$`varID#`[i]]
  naCol$varType[i] = typeof(salaryData[,naCol$`varID#`[i]])
}
naCol
```

```{r}
missingWorkClass <- salaryData[is.na(salaryData$workclass)==TRUE,]
missingWorkClass
missingOccupation <- salaryData[is.na(salaryData$occupation)==TRUE,]
missingOccupation
missingNativeCountry <- salaryData[is.na(salaryData$native_country)==TRUE,]
missingNativeCountry
missingValues <- inner_join(missingWorkClass, missingOccupation, by = "fnlwgt")
missingValues
missingValues <- inner_join(missingValues, missingNativeCountry, by = "fnlwgt")
missingValues
```

removing missing values
```{r}
salaryData = na.omit(salaryData)
length(salaryData$money)
```

# Initial EDA

Below is a basic histogram of the response variable, money
```{r}
hist <- function(variable1, variable2) {
  graph <- salaryData %>% ggplot(aes(x = variable1, fill = variable2)) + geom_bar() +facet_wrap(variable2)
  return(graph)
}
```

### Hours Per Week
```{r}
#Creating a column for grouped hours per week: 0= less than 40 hours, 1 = more than 40 hours
cbind(salaryData, "groupedHours")
salaryData$groupedHours= "<40"
salaryData$groupedHours[salaryData$hours_per_week>= 40] = ">=40"
salaryData$groupedHours = as.factor(salaryData$groupedHours)

hours<-salaryData %>% group_by(groupedHours,money) %>% summarise(cnt=n()) %>%mutate(perc=round(cnt/sum(cnt),4))%>%arrange(desc(perc))
hours[hours$money==" >50K",]
Graph <- ggplot(hours[hours$money==" >50K",],aes(x=reorder(groupedHours,-perc),y=perc,colour=groupedHours))+
  geom_bar(aes(fill=groupedHours),show.legend=T,stat="identity")+
  ylab("% of Population Making Above 50K")+
  xlab("Hours Per Week")+
  ylim(0,1)
Graph
```

### Sex
Below is a plot of the percentage of each sex that makes above 50K salary. From the graph we can see a disparity between the two groups as Males seem to be almost three times as likely to make above 50K
```{r}

g2<-salaryData %>% group_by(sex,money) %>% summarise(cnt=n()) %>%mutate(perc=round(cnt/sum(cnt),4))%>%arrange(desc(perc))
  
graph <- ggplot(g2[c(3,4),],aes(x=reorder(sex,-perc),y=perc,colour=sex))+
  geom_bar(aes(fill=sex),show.legend=T,stat="identity")+
  ylab("% of Population Making Above 50K")+
  xlab("Sex")+
  ylim(0,1)
graph
```
### Education
The graph below breaks down the percentage of populations per education group that make above 50K salary. it also divides the education variable into 3 levels: 1 for no HS Grad, 2 for up to Bachelor grad, and 3 for beyond bachelors
```{r}
#Dividing Education into groups
cbind(salaryData, "educLevel")
salaryData$educLevel[salaryData$education==" Preschool" | 
                        salaryData$education==" 1st-4th" | 
                        salaryData$education==" 5th-6th" | 
                        salaryData$education==" 7th-8th" | 
                        salaryData$education==" 9th" | 
                        salaryData$education==" 10th" | 
                        salaryData$education==" 11th" | 
                        salaryData$education==" 12th"] = "<HS"

salaryData$educLevel[salaryData$education==" HS-grad" | 
                        salaryData$education==" Assoc-acdm" | 
                        salaryData$education==" Assoc-voc" | 
                        salaryData$education==" Some-college" | 
                        salaryData$education==" Bachelors"] = "HS-Bachelors"

salaryData$educLevel[salaryData$education==" Masters" | 
                        salaryData$education==" Prof-school" | 
                        salaryData$education==" Doctorate" ] = "Post-Bachelors"

salaryData$educLevel = as.factor(salaryData$educLevel)

#Education analysis before grouping
educ<-salaryData %>% group_by(education,money) %>% summarise(cnt=n()) %>%mutate(perc=round(cnt/sum(cnt),4))%>%arrange(desc(perc))
educ[educ$money==" >50K",]
graph <- ggplot(educ[educ$money==" >50K",],aes(x=reorder(education,-perc),y=perc,colour=education))+
  geom_bar(aes(fill=education),show.legend=T,stat="identity")+
  ylab("% of Population Making Above 50K")+
  xlab("Education Level")+
  ylim(0,1)
graph

#Education analysis AFTER grouping
groupEduc<-salaryData %>% group_by(educLevel,money) %>% summarise(cnt=n()) %>%mutate(perc=round(cnt/sum(cnt),4))%>%arrange(desc(perc))
groupEduc[groupEduc$money==" >50K",]
groupGraph <- ggplot(groupEduc[groupEduc$money==" >50K",],aes(x=reorder(educLevel,-perc),y=perc,colour=educLevel))+
  geom_bar(aes(fill=educLevel),show.legend=T,stat="identity")+
  ylab("% of Population Making Above 50K")+
  xlab("Education Level (Grouped)")+
  ylim(0,1)
groupGraph
```
### Age
Age histogram
```{r}
hist(salaryData$age, salaryData$money)
```
### country
```{r}
g2<-salaryData %>% group_by(native_country,money) %>% summarise(cnt=n()) %>%mutate(perc=round(cnt/sum(cnt),4))%>%arrange(desc(perc))
g2
graph <- ggplot(g2[g2$money==" >50K",],aes(x=reorder(native_country,-perc),y=perc,colour=native_country))+
  geom_bar(aes(fill=native_country),show.legend=T,stat="identity")+
  ylab("% of Population Making Above 50K")+
  xlab("Native Country")+
  ylim(0,1)
graph
```
### Race
```{r}
raceData<-salaryData %>% group_by(race,money) %>% summarise(cnt=n()) %>%mutate(perc=round(cnt/sum(cnt),4))%>%arrange(desc(perc))
raceData[raceData$money==" >50K",]
raceGraph <- ggplot(raceData[raceData$money==" >50K",],aes(x=reorder(race,-perc),y=perc,colour=race))+
  geom_bar(aes(fill=race),show.legend=T,stat="identity")+
  ylab("% of Population Making Above 50K")+
  xlab("Race")+
  ylim(0,1)
raceGraph
```
From the plot above, there is evidence to suggest that we can group Asian/Pac-Islander and White together as one and combine the other 3 groups into a second
```{r}
#0 represents other races, 1 represents asian/pac-islander/white
cbind(salaryData, "groupedRace")
salaryData$groupedRace = "Other"
salaryData$groupedRace[salaryData$race==" Asian-Pac-Islander" | salaryData$race==" White"] = "Asian/PacIsland/White"
salaryData$groupedRace = as.factor(salaryData$groupedRace)

raceData<-salaryData %>% group_by(groupedRace,money) %>% summarise(cnt=n()) %>%mutate(perc=round(cnt/sum(cnt),4))%>%arrange(desc(perc))
raceData[raceData$money==" >50K",]
raceGraph <- ggplot(raceData[raceData$money==" >50K",],aes(x=reorder(groupedRace,-perc),y=perc,colour=groupedRace))+
  geom_bar(aes(fill=groupedRace),show.legend=T,stat="identity")+
  ylab("% of Population Making Above 50K")+
  xlab("Race")+
  ylim(0,1)
raceGraph
```

Capital gain plot: Definitely shows signs of a relationship between capital gain and making above 50K
```{r}
ggplot(data=salaryData, aes(x = fnlwgt, y = capital_gain, col=money))+geom_point()
```

###Marriage
```{r}
maritalStatus<-salaryData %>% group_by(marital_status,money) %>% summarise(cnt=n()) %>%mutate(perc=round(cnt/sum(cnt),4))%>%arrange(desc(perc))
maritalStatus[maritalStatus$money==" >50K",]
Graph <- ggplot(maritalStatus[maritalStatus$money==" >50K",],aes(x=reorder(marital_status,-perc),y=perc,colour=marital_status))+
  geom_bar(aes(fill=marital_status),show.legend=T,stat="identity")+
  ylab("% of Population Making Above 50K")+
  xlab("Education Level (Grouped)")+
  ylim(0,1)
Graph
```
We can separate marriage status out into 2 groups: active marriage and non-active

```{r}
#Creating a column for grouped marital status: 0=Not married/non-active marriage/divorce, 1 = actively married
cbind(salaryData, "groupedMarital")
salaryData$groupedMarital= "Other"
salaryData$groupedMarital[salaryData$marital_status==" Married-civ-spouse"|salaryData$marital_status==" Married-AF-spouse"] = "Actively Married"
salaryData$groupedMarital = as.factor(salaryData$groupedMarital)

maritalStatus<-salaryData %>% group_by(groupedMarital,money) %>% summarise(cnt=n()) %>%mutate(perc=round(cnt/sum(cnt),4))%>%arrange(desc(perc))
maritalStatus[maritalStatus$money==" >50K",]
Graph <- ggplot(maritalStatus[maritalStatus$money==" >50K",],aes(x=reorder(groupedMarital,-perc),y=perc,colour=groupedMarital))+
  geom_bar(aes(fill=groupedMarital),show.legend=T,stat="identity")+
  ylab("% of Population Making Above 50K")+
  xlab("Marital Status")+
  ylim(0,1)
Graph
```

```{r}
g2<-salaryData %>% group_by(relationship,money) %>% summarise(cnt=n()) %>%mutate(perc=round(cnt/sum(cnt),4))%>%arrange(desc(perc))
  
graph <- ggplot(g2[g2$money==" >50K",],aes(x=reorder(relationship,-perc),y=perc,colour=relationship))+
  geom_bar(aes(fill=relationship),show.legend=T,stat="identity")+
  ylab("% of Population Making Above 50K")+
  xlab("Sex")+
  ylim(0,1)
graph
```

```{r}
#Creating a column for grouped relationship status: 0=other, 1 = married
cbind(salaryData, "groupedRelationsip")
salaryData$groupedRelationship= "Other"
salaryData$groupedRelationship[salaryData$relationship==" Wife"|salaryData$relationship==" Husband"] = "Married"
salaryData$groupedRelationship = as.factor(salaryData$groupedRelationship)

relation<-salaryData %>% group_by(groupedRelationship,money) %>% summarise(cnt=n()) %>%mutate(perc=round(cnt/sum(cnt),4))%>%arrange(desc(perc))
relation[relation$money==" >50K",]
Graph <- ggplot(relation[relation$money==" >50K",],aes(x=reorder(groupedRelationship,-perc),y=perc,colour=groupedRelationship))+
  geom_bar(aes(fill=groupedRelationship),show.legend=T,stat="identity")+
  ylab("% of Population Making Above 50K")+
  xlab("Marital Status")+
  ylim(0,1)
Graph
```

### Work Class
```{r}
work<-salaryData %>% group_by(workclass,money) %>% summarise(cnt=n()) %>%mutate(perc=round(cnt/sum(cnt),4))%>%arrange(desc(perc))
work[work$money==" >50K",]
Graph <- ggplot(work[work$money==" >50K",],aes(x=reorder(workclass,-perc),y=perc,colour=workclass))+
  geom_bar(aes(fill=workclass),show.legend=T,stat="identity")+
  ylab("% of Population Making Above 50K")+
  xlab("Work Class (Grouped)")
Graph
```
a majority of self employed incorporations make above 50k a year

### Occupation
```{r}
occ<-salaryData %>% group_by(occupation,money) %>% summarise(cnt=n()) %>%mutate(perc=round(cnt/sum(cnt),4))%>%arrange(desc(perc))
occ[occ$money==" >50K",]
Graph <- ggplot(occ[occ$money==" >50K",],aes(x=reorder(occupation,-perc),y=perc,colour=occupation))+
  geom_bar(aes(fill=occupation),show.legend=T,stat="identity")+
  ylab("% of Population Making Above 50K")+
  xlab("Education Level (Grouped)")+
  ylim(0,1)
Graph
```


Creating correlation plots for all continuous variables while also including the response variable
```{r}
continuousVar = salaryData[,c(1,3,5,11,12,13)]
panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- abs(cor(x, y))
    txt <- format(c(r, 0.123456789), digits = digits)[1]
    txt <- paste0(prefix, txt)
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt, cex = cex.cor * r)
}
pairs(continuousVar, lower.panel = panel.smooth, upper.panel = panel.cor,
      gap=0, row1attop=FALSE)
```

From the plot above, we can see that there are no concerns for colinearity between any of the continuous variables.

### Final Data Reduction
```{r}
colnames(salaryData)
groupedVarsWithCountry = salaryData[,c(15,1,2,3,5,7,10,11,12,16,17,18,19,20,14)]
groupedVars = salaryData[,c(15,1,2,3,5,7,10,11,12,16,17,18,19,20)]
head(groupedVars)
```

# Objective 1
## Logistic Model Building
### Data Split
Train Test Split using dataset containing reduced factor levels
```{r}
set.seed(77)
split = .85
index<-sample(1:dim(groupedVars)[1],round(split * dim(groupedVars)[1]))
test<-groupedVars[-index,]
train<-groupedVars[index,]
str(train)
```

### Feature selection on full model
running lasso variable selection
```{r}
dat.train.x <- model.matrix(money~.,train)
dat.train.y<-train[,1]
cvfit <- cv.glmnet(dat.train.x, dat.train.y, family = "binomial", type.measure = "class", nlambda = 1000)
plot(cvfit)
coef(cvfit, s = "lambda.min")
#CV misclassification error rate is little below .1
print("CV Error Rate:")
cvfit$cvm[which(cvfit$lambda==cvfit$lambda.min)]

#Optimal penalty
print("Penalty Value:")
cvfit$lambda.min

#For final model predictions go ahead and refit lasso using entire
#data set
finalmodel<-glmnet(dat.train.x, dat.train.y, family = "binomial",lambda=cvfit$lambda.min)
```

stepwise
```{r}
full.log<-glm(money~.,family="binomial",data=train)
step.log<-full.log %>% stepAIC(trace=FALSE)
```

```{r}
back.log <- full.log %>% stepAIC(trace=FALSE, direction = "backward")
summary(back.log)
```
from the breakdown of the parameters above, we can see that the following variables do not have much significance to the model: Capital Gain, Capital Loss, FnlWgt. We can also see that occupation has a lower percentage of levels that are statistically insignificant when compared with the Work Class variable.

```{r}
dat.test.x<-model.matrix(money~.,test)
fit.pred.lasso <- predict(finalmodel, newx = dat.test.x, type = "response")

test$money[1:15]
fit.pred.lasso[1:15]

#Making predictions for stepwise as well for later
fit.pred.step<-predict(step.log,newdata=test,type="response")
```

```{r}
#Lets use the predicted probablities to classify the observations and make a final confusion matrix for the two models.  We can use it to calculate error metrics.
#Lets use a cutoff of 0.5 to make the classification.
cutoff<-0.5
class.lasso<-factor(ifelse(fit.pred.lasso>cutoff,">50K","<=50K"),levels=c("<=50K",">50K"))
class.step<-factor(ifelse(fit.pred.step>cutoff,">50K","<=50K"),levels=c("<=50K",">50K"))

#Confusion Matrix for Lasso
conf.lasso<-table(class.lasso,test$money)
print("Confusion matrix for LASSO")
conf.lasso

conf.step<-table(class.step,test$money)
print("Confusion matrix for Stepwise")
conf.step
```

```{r}
#Accuracy of LASSO and Stepwise
print("Overall accuracy for LASSO and Stepwise respectively")
sum(diag(conf.lasso))/sum(conf.lasso)
sum(diag(conf.step))/sum(conf.step)


print("Alternative calculations of accuracy")
#Rather than making the calculations from the table, we can compute them more quickly using the following code which just checks if the prediction matches the truth and then computes the proportion.
mean(class.lasso==test$money)
mean(class.step==test$money)
```

```{r}
library(ROCR)
results.lasso<-prediction(fit.pred.lasso, test$money,label.ordering=c(" <=50K"," >50K"))
roc.lasso = performance(results.lasso, measure = "tpr", x.measure = "fpr")
plot(roc.lasso,colorize = TRUE)
abline(a=0, b= 1)
```
### Final Model
```{r}
#Creating X and Y variables within model
trainX <- model.matrix(money~age + sex + educLevel + groupedHours + occupation + groupedRelationship,train)
trainY<-train[,1]

#Creating model
model<-glmnet(trainX, trainY, family = "binomial",lambda=cvfit$lambda.min)
testX<-model.matrix(money~age + sex + educLevel + groupedHours + occupation + groupedRelationship ,test)

#Doing Predictions
predictions <- predict(model, newx = testX, type = "response")
cutoff<-0.5
customModel<-factor(ifelse(predictions>cutoff,">50K","<=50K"),levels=c("<=50K",">50K"))

#Confusion Matrix
customConfInt<-table(customModel,test$money)
print("Confusion matrix for Custom Model")
customConfInt

#Accuracy of custom model
print("Overall accuracy for custom model")
sum(diag(customConfInt))/sum(customConfInt)
results<-prediction(predictions, test$money,label.ordering=c(" <=50K"," >50K"))

#ROC Curve
roc = performance(results, measure = "tpr", x.measure = "fpr")
plot(roc,colorize = TRUE)
abline(a=0, b= 1)
```
# Objective 2
## Complex Model
### LDA
```{r}
str(groupedVars)
mylda <- lda(money ~ ., data = groupedVars[-3])
prd <- as.numeric(predict(mylda, newdata = nd)$class)
```