---
title: "MSDS-6372-Project-2"
author: "Garret Shankel, Chad Kwong, Ryan Herrin"
date: "7/17/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Libs
library(stringr)
library(plotly)
library(ggplot2)
library(plyr) 
library(dplyr)
library(tidyr)
```

# Predicting if someone makes over 50k

## Two Objectives of the project 
+ Objective 1: Display the ability to perform EDA and build a logisitc regression model. 
+ Objective 2:  With a simple logistic regression model as a baseline, perform additional competing \
models to improve on prediction performance metrics.  Which metrics to compare models and evaluate \
performance are up to you and your given data set.

#### Import data
After data is assigned and headers added, these data frames should not be modified to keep 
integrity. Different versions should be made using them instead. 
```{r}
# Locations of data files. Extensions are different but all are in CSV format 
adult_loc <- 'data/adult.data'
adult_test_loc <- 'data/adult.test'
# Read in data
adult_data <- read.csv(adult_loc, sep=',', header = FALSE)
adult_test_data <- read.csv(adult_test_loc, sep=',', header = FALSE) 

# Define and set column names 
adult_names <- c('age', 'workclass', 'fnlwgt', 'education', 'education-num', 'marital-status', 
                 'occupation', 'relationship', 'race', 'sex', 'capital-gain', 'capital-loss',
                 'hours-per-week', 'native-country')

colnames(adult_data) <- adult_names
colnames(adult_test_data) <- adult_names
```






















