---
title: "MSDS-6372-Project-2"
author: "Garret Shankel, Chad Kwong, Ryan Herrin"
date: "7/17/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Libs
library(stringr)
library(plotly)
library(ggplot2)
library(plyr) 
library(dplyr)
library(tidyr)
```

# Predicting if someone makes over 50k

## Two Objectives of the project 
+ Objective 1: Display the ability to perform EDA and build a logisitc regression model. 
+ Objective 2:  With a simple logistic regression model as a baseline, perform additional competing \
models to improve on prediction performance metrics.  Which metrics to compare models and evaluate \
performance are up to you and your given data set.

#### Import data
After data is assigned and headers added, these data frames should not be modified to keep 
integrity. Different versions should be made using them instead. 
```{r}
# Locations of data files. Extensions are different but all are in CSV format 
adult_loc <- 'data/adult.data'
adult_test_loc <- 'data/adult.test'
# Read in data
adult_data <- read.csv(adult_loc, sep=',', header = FALSE)
adult_test_data <- read.csv(adult_test_loc, sep=',', header = FALSE) 

# Define and set column names 
adult_names <- c('age', 'workclass', 'fnlwgt', 'education', 'education_num', 'marital_status', 
                 'occupation', 'relationship', 'race', 'sex', 'capital_gain', 'capital_loss',
                 'hours_per_week', 'native_country', 'money')

colnames(adult_data) <- adult_names
colnames(adult_test_data) <- adult_names
```

#### Data Delta and Cleaning 
```{r}
# Create new Data Frames.
adult <- adult_data
adult_test <- adult_test_data

# Create one data set to make our own train/test split 
adult_combined <- rbind(adult, adult_test)
```

Function to apply same changes accross all data frames
```{r, include=FALSE}
apply_mods <- function(df) {
  # Function to apply all modifications to a data frame and return it back
  delta_df <- df 
  
  # Replace all question marks with NA values 
  delta_df[delta_df == " ?"] <- "NA"
  
  # There should only be two factors for money. Need to fix that. 
  delta_df[delta_df == " <=50K."] <- "<=50K"
  delta_df[delta_df == " >50K."] <- ">50K"
  delta_df[delta_df == " <=50K"] <- "<=50K"
  delta_df[delta_df == " >50K"] <- ">50K"
  
  # transform data type "chr" columns into "factor"
  delta_df <- transform(delta_df, 
                        workclass=as.factor(workclass),
                        education_num=as.factor(education_num),
                        marital_status=as.factor(marital_status),
                        occupation=as.factor(occupation),
                        relationship=as.factor(relationship),
                        race=as.factor(race),
                        sex=as.factor(sex),
                        native_country=as.factor(native_country),
                        money=as.factor(money))
  
  return(delta_df)
}
```

Datasets to apply function too
```{r}
# Apply the mods funtion to the dataframes below 
adult_combined <- apply_mods(adult_combined)

# Print Summary of combined
str(adult_combined)
```

#### EDA - R.H.
```{r, fig.show="hold", out.width="50%"}
# Age
ggplot(adult_combined, aes(age, colour=money)) + geom_density()
```




















